

```{r init}
opts_chunk$set(cache.path="B:/Conservation_Limits/CL_Juvenile_Density/Spatial_Abundance_Model/package/CLdata/data-raw/cache/ManipulateRawData/",
  fig.path="B:/Conservation_Limits/CL_Juvenile_Density/Spatial_Abundance_Model/package/CLdata/data-raw/fig/ManipulateRawData/",
  tidy = FALSE)
```



# Maniplate raw data

This chapter summarises the manipulation required of the available raw data.  Removal of unnessisary data and connecting of data sets.


##### Outline [to be re-writen]

* [add river segment and catchment ID to site info](#site-info) Three projections are set up: BNG (Brittish National Grid) in Eastings and Northings; WGS84 (GS 1984) in lattitude and longitude; and Mercator.  All data is plotted in BNG, however, if set against a map downloaded from google maps then the projection is WGS84.
  
* [add barriers to site info](#site-bars) Catchment and regional data sets are associated with spatial polygon shape files read in using the `rgdal` package. There are several data sets available: catchments > 25 m2; catchments < 25 m2 (these are generally coastal catchments); hydrometric areas; subcatchments (where large catchments are subdivided); various salmon administrative regions such as the salmon statistical areas and the 1867 salmon region definitions.
  
* [calculate route to sea for barriers](#sea-route) There is one single river dataset assiciated with a measurale spatial lines shape file. 

* [trim rivers and remove psuedonodes](#trim-rivs) There is a comprehensive spatial points dataset containing barriers to fish movement.

The directories containing the raw data and where raw compressed versions are saved are:

```{r data_dir, cache=TRUE}
  # where intermediate working data is saved
  raw_dir <- "B:/Conservation_Limits/CL_Juvenile_Density/Spatial_Abundance_Model/package/CLdata/data-raw/rData/"

  # where raw data resides
  data_dir <- "B:/Conservation_Limits/CL_Juvenile_Density/"

  # helper fuction to get file names 
  fname <- function(file, ext = ".rData", dir = raw_dir) {
    paste0(dir, file, ext)
  }
```

<!-- 
_______________________________________________________
_______________________________________________________
*******************************************************
-->


The following functions can be used to navigate from a point on a river network to the river mouth.  As long as the lines end points match in space then once a line segment is has no onward connection, the end point is defined as a river mouth.

```{r rivernext, cache=TRUE}
# some useful functions
riverNext <- function(ID) {
  L1 <- rivs[rivs $ OBJECTID == ID,]
  if (L1 $ TNODE_ == 0) {
    cat("unconnected node\n")
    return(NA)
  }
  if (sum(rivs $ FNODE_ == L1 $ TNODE_)==0) {
    #cat("river mouth\n")
    return(NA)
  }
  L2s <- rivs[rivs $ FNODE_ == L1 $ TNODE_,]

  flocs <- sapply(L2s @ lines, function(ll) ll @ Lines[[1]] @ coords[1,])
  tloc <- tail(L1 @ lines[[1]] @ Lines[[1]] @ coords, 1)
  dist <- sqrt(colSums((flocs - c(tloc))^2))
  if (all(dist > 10))  {
    #cat("river mouth\n")
    return(NA)
  }
  L2ind <- which.min(dist) 
  L2 <- L2s[L2ind,]

  rbind(L1, L2)
}

riverEnd <- function(ID, add = FALSE, old = NULL) {
  route <- rep(NA, 100)
  route[1] <- ID
  rnext <- riverNext(ID)
  i <- 1
  while(!identical(rnext, NA) & !(route[i] %in% old)) {
    route[i <- i + 1] <- rnext $ OBJECTID[2]
    rnext <- riverNext(route[i])
  }
  route[!is.na(route)]
}
```

<!-- 
_______________________________________________________
_______________________________________________________
-->


<h2><a name="site-info">Add river segment and catchment ID to site info</a></h2>


```{r gis_stuff, cache=TRUE, dependson=c("data_dir", "rivernext")}
# get all segements with a start and end within a bbox
load(fname("ctm"))
load(fname("cctm"))
load(fname("coast"))
load(fname("rivs"))
load(fname("gis"))

#find centre of cctm polygons
ccentre <- sapply(cctm @ polygons, function(x) colMeans(x @ Polygons[[1]] @ coords))

# get river node locations
floc <- sapply(rivs @ lines, function(ll) ll @ Lines[[1]] @ coords[1,])
tloc <- sapply(rivs @ lines, function(ll) {x <- ll @ Lines[[1]] @ coords; x[nrow(x),]})

# create container for river stretch ID
gis $ rivs_OBJECTID <- NA
gis $ CATCH_ID <- NA
gis $ DESCRIPTIO <- NA
rivs $ CATCH_ID <- NA

routeList <- vector("list", nrow(gis))
names(routeList) <- 1:nrow(gis)

score <- rep(NA, nrow(gis))
visualise <- FALSE
verbose <- FALSE
tries <- order(gis $ Distance_s, decreasing = TRUE)
require(rgeos)

j <- 0
for (i in tries) {
  j <- j + 1

  if (verbose) cat("       \r", sum(is.na(gis $ CATCH_ID)), "to go!"); flush.console()
  # choose a point
  wkgis <- gis[i,]

  # take a circle round the point
  p <- c(wkgis @ coords)
  dist <- 5000
  rwhich <- which(colSums((floc - p)^2) < dist^2 | colSums((tloc - p)^2) < dist^2)
  wkrivs <- rivs[rwhich,]
  cwhich <- which(colSums((ccentre - p)^2) < dist^2)
  if (length(cwhich) == 0) cwhich = 1 # quick lazy fix
  wkcctm <- cctm[cwhich,]
  if (visualise) {
    plot(wkrivs, main = i)
    plot(wkcctm, add = TRUE, border = "darkred", col = "lightyellow")
    plot(ctm, add = TRUE, col = "lightyellow3")
    plot(wkrivs, col = "blue", add = TRUE)
    points(wkgis, col = "red", pch = 16)
    points(gis, col = "red")
  }  
  # find which river segment it is on
  wkdf <- 
  do.call(rbind, lapply(1:nrow(wkrivs),
         function(i) {
           lls <- wkrivs @ lines [[i]] @ Lines[[1]] @ coords
           colnames(lls) <- c("x", "y")
           lls <- as.data.frame(lls)
           lls $ OBJECTID = wkrivs $ OBJECTID[i]
           lls
         }))

  ind <- which.min((wkdf $ x - p[1])^2 + (wkdf $ y - p[2])^2)
  rind <- which(wkrivs $ OBJECTID == wkdf $ OBJECTID[ind])
  
  gis $ rivs_OBJECTID[i] <- wkrivs $ OBJECTID[rind]

  if (gis $ rivs_OBJECTID[i] %in% unlist(routeList)) {
    init <- sapply(routeList, function(x) if (length(x)) any(x %in% gis $ rivs_OBJECTID[i]) else FALSE)
    which <- which(init)[1]
    ii <- which(routeList[[which]] == gis $ rivs_OBJECTID[i])
    routeIDs <- routeList[[which]][ii : length(routeList[[which]])]

    routeList[[i]] <- routeIDs
    gis $ CATCH_ID[i] <- gis $ CATCH_ID[which]    
    gis $ DESCRIPTIO[i] <- gis $ DESCRIPTIO[which]    
    score[i] <- score[which]

    # add info to rivs
    rivs $ CATCH_ID[rivs $ OBJECTID %in% routeIDs] <- gis $ CATCH_ID[which]

    rll <- gLineMerge(rivs[rivs $ OBJECTID %in% routeIDs,])     

    if (visualise) {
      plot(rll, lwd = 10, add = TRUE)
      plot(rll, col = "orange", lwd = 5, add = TRUE)
    }

    next
  }
  
  # find route to sea and choose catchment
  routeIDs <- riverEnd(gis $ rivs_OBJECTID[i], old = unlist(routeList))

  if (any(routeIDs %in% unlist(routeList))) {
    init <- sapply(routeList, function(x) if (length(x)) any(x %in% routeIDs) else FALSE)
    which <- which(init)[1]

    ii <- which(routeList[[which]] %in% routeIDs)
    routeIDs <- c(routeIDs, routeList[[which]][ii : length(routeList[[which]])][-1])

    routeList[[i]] <- routeIDs
    gis $ CATCH_ID[i] <- gis $ CATCH_ID[which]    
    gis $ DESCRIPTIO[i] <- gis $ DESCRIPTIO[which]    
    score[i] <- score[which]

    # add info to rivs
    rivs $ CATCH_ID[rivs $ OBJECTID %in% routeIDs] <- gis $ CATCH_ID[which]

    rll <- gLineMerge(rivs[rivs $ OBJECTID %in% routeIDs,])     

    if (visualise) {
      plot(rll, lwd = 10, add = TRUE)
      plot(rll, col = "orange", lwd = 5, add = TRUE)
    }

    next
  }

  rll <- gLineMerge(rivs[rivs $ OBJECTID %in% routeIDs,])     

  if (visualise) {
    plot(rll, lwd = 10, add = TRUE)
    plot(rll, col = "orange", lwd = 5, add = TRUE)
  }

  routeList[[i]] <- routeIDs

  # find which catchment polygon the majority of nodes lie in.
  # this is nessisaru becouse some rivers have nodes in more than
  # one catchment polygon

  member1 <- 
    sapply(wkcctm @ polygons, function(x)
      mean(point.in.polygon(rll @ lines[[1]] @ Lines[[1]] @ coords[,1], 
                            rll @ lines[[1]] @ Lines[[1]] @ coords[,2], 
                            x @ Polygons[[1]] @ coords[,1],
                            x @ Polygons[[1]] @ coords[,2])))

  member2 <- 
    sapply(ctm @ polygons, function(x)
      mean(point.in.polygon(rll @ lines[[1]] @ Lines[[1]] @ coords[,1], 
                            rll @ lines[[1]] @ Lines[[1]] @ coords[,2], 
                            x @ Polygons[[1]] @ coords[,1],
                            x @ Polygons[[1]] @ coords[,2])))

  if (max(member1) > max(member2)) {
    cpoly <- wkcctm[which.max(member1),]
  } else {
    cpoly <- ctm[which.max(member2),]
  }

  if (visualise) {
    plot(cpoly, add = TRUE, lwd = 3)
    points(wkgis, col = "red", pch = 16)
    points(gis, col = "red")
  }      

  gis $ CATCH_ID[i] <- cpoly $ CATCH_ID    
  gis $ DESCRIPTIO[i] <- as.character(cpoly $ DESCRIPTIO)    

  # add info to rivs
  rivs $ CATCH_ID[rivs $ OBJECTID %in% routeIDs] <- cpoly $ CATCH_ID

  # store the score: the proportion of nodes in the allocated
  # polygon
  score[i] <- max(member1, member2)
}

# save reduced rivers dataset to rData folder
redrivs <- rivs[!is.na(rivs $ CATCH_ID),]
save(redrivs, file = fname("redrivs"))

# save routes to sea dataset to rData folder
save(routeList, file = fname("routeList"))

# save rediced catchments dataset to rData folder
redctm1 <- ctm[ctm $ CATCH_ID %in% gis $ CATCH_ID,]
redctm1 @ data <- redctm1 @ data[c("CATCH_ID", "DESCRIPTIO")]
redctm1 $ COASTAL <- FALSE

redctm2 <- cctm[cctm $ CATCH_ID %in% gis $ CATCH_ID,]
redctm2 @ data <- redctm2 @ data[c("CATCH_ID", "DESCRIPTIO")]
redctm2 $ COASTAL <- TRUE

redctm <- rbind(redctm1, redctm2)
save(redctm, file = fname("redctm"))
```

<!-- 
_______________________________________________________
_______________________________________________________
-->


<h2><a name="site-bars">add barriers to site info</a></h2>

```{r redbars, cache=TRUE, depenson=c("data_dir", "gis_stuff")}

# now need to link bars with catchment
load(fname("bars"))
wk <- over(bars, redctm) 
bars @ data <- cbind(bars @ data, wk)
bars <- bars[!is.na(bars $ CATCH_ID),]

# and match to river segment
load(fname("rivs"))
floc <- sapply(rivs @ lines, function(ll) ll @ Lines[[1]] @ coords[1,])
tloc <- sapply(rivs @ lines, function(ll) {x <- ll @ Lines[[1]] @ coords; x[nrow(x),]})

# create container for river stretch ID
bars $ rivs_OBJECTID <- NA
# find river segment  
verbose <- FALSE
for (i in 1:nrow(bars)) {
  if (verbose) cat("       \r", sum(is.na(bars $ rivs_OBJECTID)), "to go!"); flush.console()
  # choose a point
  wkbars <- bars[i,]

  # take a circle round the point
  p <- c(wkbars @ coords)
  dist <- 5000
  rwhich <- which(colSums((floc - p)^2) < dist^2 | colSums((tloc - p)^2) < dist^2)
  wkrivs <- rivs[rwhich,]
  # find which river segment it is on
  wkdf <- 
  do.call(rbind, lapply(1:nrow(wkrivs),
         function(i) {
           lls <- wkrivs @ lines [[i]] @ Lines[[1]] @ coords
           colnames(lls) <- c("x", "y")
           lls <- as.data.frame(lls)
           lls $ OBJECTID = wkrivs $ OBJECTID[i]
           lls
         }))

  ind <- which.min((wkdf $ x - p[1])^2 + (wkdf $ y - p[2])^2)
  rind <- which(wkrivs $ OBJECTID == wkdf $ OBJECTID[ind])
  
  #lines(wkrivs[rind,], col = "blue", lwd = 2)
  bars $ rivs_OBJECTID[i] <- wkrivs $ OBJECTID[rind]
}

# remove barriers not on our river system
which <- which(bars $ rivs_OBJECTID %in% redrivs $ OBJECTID)
redbars <- bars[which,]

save(redbars, file = fname("redbars"))
```

<!-- 
_______________________________________________________
_______________________________________________________
-->

<h2><a name="sea-route">Calculate route to sea for barriers</a></h2>


```{r barslist, eval=TRUE, cache=TRUE, dependson=c("data_dir", "gis_stuff", "redbars")}
verbose <- FALSE
visualise <- FALSE
barriersList <- 
  lapply(1:nrow(gis), function(i) {
  # choose a catchment
    if (verbose) cat("       \r", i); flush.console()
    wkgis <- gis[i,]  

    rwhich <- sapply(routeList[[i]], function(x) which(redrivs $ OBJECTID == x))
    wkrivs <- redrivs[rwhich,]
    # need to find route to sea and list coords
    rline <- do.call(rbind, lapply(wkrivs @ lines, function(x) x @ Lines[[1]] @ coords))
    rline <- unique(rline)
    nodetype <- rep(1, nrow(rline))

    # insert barriers
    bwhich <- which(bars $ rivs_OBJECTID %in% routeList[[i]])
    if (length(bwhich) == 0) return(character(0)) 
    wkbars <- bars[bwhich,]  

    bcoords <- snapPointsToLines(wkbars, wkrivs) @ coords

    for(j in 1:nrow(bcoords)) {
      which <- min(order(colSums((t(rline) - c(bcoords[j,]))^2))[1:2])
      rline <- rbind(rline[1:which,], bcoords[j,], rline[(which + 1):nrow(rline),])
      nodetype <- c(nodetype[1:which], paste(wkbars $ ID[j]), nodetype[(which + 1):length(nodetype)])
    }

    # now add in sample point
    gcoords <- snapPointsToLines(wkgis, wkrivs) @ coords

    # insert points
    which <- min(order(colSums((t(rline) - c(gcoords))^2))[1:2])
    rline <- rbind(rline[1:which,], gcoords, rline[(which + 1):nrow(rline),])
    nodetype <- c(nodetype[1:which], "gis", nodetype[(which + 1):length(nodetype)])

    if (visualise) {
      plot(rline, col = "blue", pch = 16, type = "b")
      points(rline[nodetype == "gis",,drop=FALSE], col = "red", pch = 16)
      points(rline[nodetype != "1" & nodetype != "gis",,drop=FALSE], col = 1, pch = 4, lwd = 3)
      points(rline[nrow(rline),,drop=FALSE], col = "green", pch = 16)
    }

    gbars <- nodetype[(which(nodetype=="gis")+1):length(nodetype)]
    gbars[gbars != "1"]
  })


  # save barriers dataset to rData folder
  save(barriersList, file = fname("barriersList"))

  # add barrier info to gis
  impass <- bars $ BARRIER_PE %in% c("Impassable natural waterfall or other natural obstruction",
                                      "Impassable obstructions Type unknown.",
                                      "Impassable weir, dam or other manmade structure") 
  gis $ barrier <- 
    sapply(barriersList, function(x) {
      if(length(x)) {
        any(impass[bars $ ID %in% as.numeric(x)])
      } else {
        FALSE
      } 
    })

  # finally save all changes to gis dataset to rData folder
  save(gis, file = fname("gis_v2"))
```


<!-- 
_______________________________________________________
_______________________________________________________
-->

<h2><a name="trim-rivs">Trim rivers and remove psuedonodes</a></h2>

```{r trim, cache=TRUE, dependson=c("data_dir", "barslist","sea-route")}
# discard nonsampled river branches
catchments <- sort(unique(gis $ CATCH_ID))
newrivs <- list(length(catchments))

verbose <- FALSE
visualise <- FALSE
for (catch in catchments) {
  if (verbose) cat("\ncleaning catchment:", catch[1], "-", 
      as.character(subset(gis, CATCH_ID %in% catch) $ DESCRIPTIO[1]), 
      "\n"); flush.console()
  
  wkgis <- gis[gis $ CATCH_ID %in% catch,]
  wkrivs <- redrivs[redrivs $ CATCH_ID %in% catch,]
  if (visualise) {
    plot(redctm[which(redctm $ CATCH_ID %in% catch),], col = "lightyellow")
    title(main = paste0(catch[1], ": ", subset(gis, CATCH_ID %in% catch) $ DESCRIPTIO[1]))
    plot(wkrivs, col = grey(0.5), add = TRUE)
    points(wkgis, col = "red")
  }
  # go through each site, map route to sea recording nodes on route
  # to save time a route can be terminated if the node has been seen before

  # make river into a dataframe
  wkdf <- 
    do.call(rbind, lapply(1:nrow(wkrivs),
           function(i) {
             lls <- wkrivs @ lines [[i]] @ Lines[[1]] @ coords
             colnames(lls) <- c("x", "y")
             lls <- as.data.frame(lls)
             lls $ OBJECTID = wkrivs $ OBJECTID[i]
             lls
           }))

  ## now to remove psuedonodes
  wk <- wkrivs#[wkrivs $ OBJECTID %in% IDs,]
  wk $ keep <- TRUE

  i <- 0
  require(rgeos)
  while(i < nrow(wk)) {
    # pick a NODE
    node <- wk $ FNODE_[i <- i + 1]

    # if already marked for removal escape
    if (!wk $ keep[i]) next

    # what is the in degree?
    idegree <- sum(wk $ TNODE_ == node)
    odegree <- sum(wk $ FNODE_ == node)

    # early escape
    if (idegree != 1 || odegree != 1) next

    if (verbose) cat("                             \r",
            i, "psuedonode - REMOVE"); flush.console()

    oldlines <- wk[wk $ TNODE_ == node | wk $ FNODE_ == node,]
    fix <- which(wk $ TNODE_ == node)
    
    # signal for removal and loop on itself (this might not be nesisary)
    wk $ keep[wk $ FNODE_ == node] <- FALSE
    wk $ TNODE_[wk $ FNODE_ == node] <- node

    wk $ LENGTH[fix] <- sum(oldlines $ LENGTH)
    wk $ Shape_Leng[fix] <- sum(oldlines $ Shape_Leng)
    wk $ TNODE_[fix] <- oldlines $ TNODE_[oldlines $ FNODE_ == node]
    wk @ lines[[fix]] <- gLineMerge(oldlines) @ lines[[1]]

    if (FALSE) {
      plot(wk[fix,], col = "blue", lwd = 5)
      plot(oldlines, col = c("red", "green"), add = TRUE)
    }
  }

  # keep only good nodes
  wk <- wk[wk $ keep,]
  #plot(wk, lwd = 4, col = 2:10, add = TRUE)

  # save river
  newrivs[[which(catchments == catch)]] <- wk
}

## convert newrivs into a spatial lines DF
dat <- do.call(rbind, lapply(newrivs, function(x) x @ data))
Ls <- lapply(newrivs, function(x) x @ lines)
Ls <- do.call(c, Ls)
for (i in 1:nrow(dat)) Ls[[i]] @ ID <- rownames(dat)[i] 
Ls <- SpatialLines(Ls)
newrivs <- SpatialLinesDataFrame(Ls, dat)

# save full dataset to rData folder
save(newrivs, file = fname("newrivs"))
```

The reduced rivers and catchements look like

```{r plot_newrivs, cache=TRUE, dependson=c("trim-rivs")}
# plot
plot(newrivs, main = "ctm", col = grey(0.7))
```
