

```{r init}
opts_chunk$set(cache.path="B:/Conservation_Limits/CL_Juvenile_Density/Spatial_Abundance_Model/package/CLdata/data-raw/cache/ReadAllData/",
  fig.path="B:/Conservation_Limits/CL_Juvenile_Density/Spatial_Abundance_Model/package/CLdata/data-raw/fig/ReadAllData/",
  tidy = FALSE)
```



# Read raw data (part I)

This chapter summarises the reading in of the available raw spatial data.

##### Outline

* [Projections](#projections) Three projections are set up: BNG (Brittish National Grid) in Eastings and Northings; WGS84 (GS 1984) in lattitude and longitude; and Mercator.  All data is plotted in BNG, however, if set against a map downloaded from google maps then the projection is WGS84.
  
* [Catchments](#catchments) Catchment and regional data sets are associated with spatial polygon shape files read in using the `rgdal` package. There are several data sets available: catchments > 25 m2; catchments < 25 m2 (these are generally coastal catchments); hydrometric areas; subcatchments (where large catchments are subdivided); various salmon administrative regions such as the salmon statistical areas and the 1867 salmon region definitions.
  
* [River data](#river-data) There is one single river dataset assiciated with a measurale spatial lines shape file. 

* [Barriers data](#barrier-data) There is a comprehensive spatial points dataset containing barriers to fish movement.

The directories containing the raw data and where raw compressed versions are saved are:

```{r data_dir, cache=TRUE}
  # where intermediate working data is saved
  raw_dir <- "B:/Conservation_Limits/CL_Juvenile_Density/Spatial_Abundance_Model/package/CLdata/data-raw/rData/"

  # where raw data resides
  data_dir <- "B:/Conservation_Limits/CL_Juvenile_Density/"

  # helper fuction to get file names 
  fname <- function(file, ext = ".rData", dir = raw_dir) {
    paste0(dir, file, ext)
  }
```

<h2><a name="projections">Projections</a></h2>

The three projections are defined as follows:

```{r proj, cache=TRUE, dependson=c("data_dir")}
proj <- list(
  wgs84 = '+proj=longlat +datum=WGS84',
  bng   = '+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.9996012717 +x_0=400000 +y_0=-100000 +ellps=airy +datum=OSGB36 +units=m +no_defs',
  merc  = "+proj=merc +a=6378137 +b=6378137 +lat_ts=0.0 +lon_0=0.0 +x_0=0.0 +y_0=0 +k=1.0 +units=m +nadgrids=@null +no_defs"
)

# save full dataset to raw data folder
save(proj, file = fname("proj"))
```


<h2><a name="catchments">Catchments and regions</a></h2>

There are a variety of region shape files available.  The following are considered in this analysis.  Currently these data are read in using the `maptools` packages, however a better option is to use the `rgdal` package.  **This should be updated when possible**.

<!-- 
_______________________________________________________
_______________________________________________________
-->

### The ctm shapefile 


```{r ctm, cache=TRUE, dependson=c("data_dir", "proj")}
# read in baseline catchments
require(maptools)
ctm <- readShapeSpatial(fname("GIS_shapefiles_for_model/Catchments/Baseline_confluence_nested_catchments", dir = data_dir, ext = ""))

# add a column defining discrete catchments for use in analysis (should maybe do this later)
ctm $ CATCH_ID <- 1:nrow(ctm)

# add projection
ctm @ proj4string <- CRS(proj $ bng) 

# save full dataset to rData folder
save(ctm, file = fname("ctm"))
```

This data looks like

```{r plot_ctm, cache=TRUE, dependson=c("ctm")}
# plot
plot(ctm, main = "ctm", col = grey(0.7))
```

<!-- 
_______________________________________________________
_______________________________________________________
-->


### The cctm shapefile 

These polygons define the small to very small coastal catchments.  They are often grouped together, but can be considered individually.

<!-- 
should really be:
  dependson=c("data_dir", "proj", "ctm")}
but I want to limit the reruning of this segment
-->

```{r cctm, cache=TRUE, dependson=c("data_dir", "proj")}
# read in baseline catchments
require(maptools)
cctm <- readShapeSpatial(fname("GIS_shapefiles_for_model/Catchments/coastal_catchments", dir = data_dir, ext = ""))

# make names consistent
names(cctm)[names(cctm) == "CATCH_ID"] <- "CATCHMENT"
names(cctm)[names(cctm) == "MIN_CATCHN"] <- "DESCRIPTIO"

# add a column for discrete catchements as for ctm
cctm $ CATCH_ID <- nrow(ctm) + 1:nrow(cctm)

# add projection
cctm @ proj4string <- CRS(proj $ bng) 

# save full dataset to rData folder
save(cctm, file = fname("cctm"))
```

This data looks like

```{r plot_cctm, cache=TRUE, dependson=c("cctm")}
# plot
plot(cctm, main = "cctm", col = grey(0.7))
```

<!-- 
_______________________________________________________
_______________________________________________________
-->


### The sctm shapefile 

These polygons define so called super-catchments.

```{r sctm, cache=TRUE, dependson=c("data_dir", "proj")}
# read in baseline catchments
require(maptools)
sctm <- readShapeSpatial(fname("GIS_shapefiles_for_model/Catchments/supercatch", dir = data_dir, ext = ""))

# add projection
sctm @ proj4string <- CRS(proj $ bng) 

# save full dataset to rData folder
save(sctm, file = fname("sctm"))
```

This data looks like

```{r plot_sctm, cache=TRUE, dependson=c("sctm")}
# plot
plot(sctm, main = "sctm", col = grey(0.7))
```

<!-- 
_______________________________________________________
_______________________________________________________
-->


### The rctm shapefile 

These polygons define so called river catchments.


```{r rctm, cache=TRUE, dependson=c("data_dir", "proj")}
# read in baseline catchments
require(maptools)
rctm <- readShapeSpatial(fname("GIS_shapefiles_for_model/Catchments/rivcatch", dir = data_dir, ext = ""))

# add projection
rctm @ proj4string <- CRS(proj $ bng) 

# save full dataset to rData folder
save(rctm, file = fname("rctm"))
```

This data looks like

```{r plot_rctm, cache=TRUE, dependson=c("rctm")}
# plot
plot(rctm, main = "rctm", col = grey(0.7))
```

<!-- 
_______________________________________________________
_______________________________________________________
-->


### The hma shapefile 

These polygons define so called hydrometric areas.


```{r hma, cache=TRUE, dependson=c("data_dir", "proj")}
# read in baseline catchments
require(maptools)
hma <- readShapeSpatial(fname("GIS_shapefiles_for_model/SEPA_hydrometric_areas/SEPA_hydrometric_areas", dir = data_dir, ext = ""))

# rename for consistency
names(hma)[names(hma) == "HA"] <- "HACode"
names(hma)[names(hma) == "NAME"] <- "HAName"

# add projection
hma @ proj4string <- CRS(proj $ bng) 

# save full dataset to rData folder
save(hma, file = fname("hma"))
```

This data looks like

```{r plot_hma, cache=TRUE, dependson=c("hma")}
# plot
plot(hma, main = "hma", col = grey(0.7))
```

<!-- 
_______________________________________________________
_______________________________________________________
-->


### The Creating a coastline from hma shapes 

An accurate coastline is usefull when plotting coastal and island catchments.

```{r coast, cache=TRUE, dependson=c("data_dir", "hma")}
# union the hma polygons
require(rgeos)
coast <- gUnaryUnion(hma)

# save full dataset to rData folder
save(coast, file = fname("coast"))
```

This data looks like

```{r plot_coast, cache=TRUE, dependson=c("coast")}
# plot
plot(coast, main = "coast", col = grey(0.7))
```

<!-- 
_______________________________________________________
_______________________________________________________
-->


### The sa1868 shapefile 

These polygons define salmon administrative boundaries.

```{r sa1868, cache=TRUE, dependson=c("data_dir", "proj")}
# read in baseline catchments
require(maptools)
sa1868 <- readShapeSpatial(fname("GIS_shapefiles_for_model/District maps data/sa_1868_1110", dir = data_dir, ext = ""))

# add projection
sa1868 @ proj4string <- CRS(proj $ bng) 

# save full dataset to rData folder
save(sa1868, file = fname("sa1868"))
```

This data looks like

```{r plot_sa1868, cache=TRUE, dependson=c("sa1868")}
# plot
plot(sa1868, main = "sa1868", col = grey(0.7))
```

<!-- 
_______________________________________________________
_______________________________________________________
-->


### The saSFD shapefile 

These polygons define salmon administrative boundaries.

```{r saSFD, cache=TRUE, dependson=c("data_dir", "proj")}
# read in baseline catchments
require(maptools)
saSFD <- readShapeSpatial(fname("GIS_shapefiles_for_model/District maps data/sa_SFD_1110", dir = data_dir, ext = ""))

# add projection
saSFD @ proj4string <- CRS(proj $ bng) 

# save full dataset to rData folder
save(saSFD, file = fname("saSFD"))
```

This data looks like

```{r plot_saSFD, cache=TRUE, dependson=c("saSFD")}
# plot
plot(saSFD, main = "saSFD", col = grey(0.7))
```

<!-- 
_______________________________________________________
_______________________________________________________
-->


### The sastatD shapefile 

These polygons define salmon statistical disctricts.

```{r sastatD, cache=TRUE, dependson=c("data_dir", "proj")}
# read in baseline catchments
require(maptools)
sastatD <- readShapeSpatial(fname("GIS_shapefiles_for_model/District maps data/sa_statD_1110", dir = data_dir, ext = ""))

# add projection
sastatD @ proj4string <- CRS(proj $ bng) 

# save full dataset to rData folder
save(sastatD, file = fname("sastatD"))
```

This data looks like

```{r plot_sastatD, cache=TRUE, dependson=c("sastatD")}
# plot
plot(sastatD, main = "sastatD", col = grey(0.7))
```

<!-- 
_______________________________________________________
_______________________________________________________
-->

### The sastatR shapefile 

These polygons define salmon statistical regions.

```{r sastatR, cache=TRUE, dependson=c("data_dir", "proj")}
# read in baseline catchments
require(maptools)
sastatR <- readShapeSpatial(fname("GIS_shapefiles_for_model/District maps data/sa_statR_1110", dir = data_dir, ext = ""))

# add projection
sastatR @ proj4string <- CRS(proj $ bng) 

# save full dataset to rData folder
save(sastatR, file = fname("sastatR"))
```

This data looks like

```{r plot_sastatR, cache=TRUE, dependson=c("sastatR")}
# plot
plot(sastatR, main = "sastatR", col = grey(0.7))
```

<!-- 
_______________________________________________________
_______________________________________________________
-->

### The satrust shapefile 

These polygons define salmon trusts.

```{r satrust, cache=TRUE, dependson=c("data_dir", "proj")}
# read in baseline catchments
require(maptools)
satrust <- readShapeSpatial(fname("GIS_shapefiles_for_model/District maps data/sa_trust_1110", dir = data_dir, ext = ""))

# add projection
satrust @ proj4string <- CRS(proj $ bng) 

# save full dataset to rData folder
save(satrust, file = fname("satrust"))
```

This data looks like

```{r plot_satrust, cache=TRUE, dependson=c("satrust")}
# plot
plot(satrust, main = "satrust", col = grey(0.7))
```

<!-- 
_______________________________________________________
_______________________________________________________
*******************************************************
-->

<h2><a name="river-data">River data</a></h2>



```{r rivs, cache=TRUE, dependson=c("data_dir", "proj")}
# read in rivers
require(maptools)
rivs <- readShapeSpatial(fname("GIS_shapefiles_for_model/rivers/gi03_rivers_050914", dir = data_dir, ext = ""))

# define coords as brittish nation grid
rivs @ proj4string <- CRS(proj $ bng) 

# rename nodes for uniqness
floc <- sapply(rivs @ lines, function(ll) ll @ Lines[[1]] @ coords[1,])
tloc <- sapply(rivs @ lines, function(ll) {x <- ll @ Lines[[1]] @ coords; x[nrow(x),]})

newfloc <- as.numeric(apply(round(floc,1)*10, 2, paste, collapse = ""))
newtloc <- as.numeric(apply(round(tloc,1)*10, 2, paste, collapse = ""))

rivs $ FNODE_ <- newfloc
rivs $ TNODE_ <- newtloc

# remove two deadend river segments getting in the way 
rivs <- rivs[rivs $ OBJECTID != 72354,] # WORKS!!
rivs <- rivs[rivs $ OBJECTID != 3887,] # WORKS!!

# save full dataset to rData folder
save(rivs, file = fname("rivs"))
```
This data looks like

```{r plot_rivs, cache=TRUE, dependson=c("rivs")}
# plot
plot(rivs, main = "rivs")
```

<!-- 
_______________________________________________________
_______________________________________________________
*******************************************************
-->

<h2><a name="barrier-data">Locations of potential barriers to fish movement</a></h2>

```{r bars, cache=TRUE, dependson=c("data_dir", "proj")}
# read in barriers
require(maptools)
bars <- readShapeSpatial(fname("GIS_shapefiles_for_model/rivers/Obstacles_to_Fish_Migration", dir = data_dir, ext = ""))

# define coords as brittish nation grid
bars @ proj4string <- CRS(proj $ bng) 

# save full dataset to rData folder
save(bars, file = fname("bars"))
```

This data looks like

```{r plot_bars, cache=TRUE, dependson=c("bars")}
# plot
plot(bars, main = "bars")
```

<!-- 
_______________________________________________________
_______________________________________________________
*******************   END     *************************
-->


